---
title: "Untitled"
format: html
editor: visual
---

## **Practice Model Selection**

First, installing and loading the packages I need.

```{r}
install.packages("naniar")
install.packages("MuMIn")
install.packages("AICcmodavg")
```

```{r}
library(tidyverse)
library(skimr)     
library(naniar)    
library(MuMIn)       
library(AICcmodavg)
```

Loading data,

Since the data file this time is a .txt file, I will use read_tsv() function in R.

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f)

```

For looking at the structure of the data,

```{r}
glimpse(d)
names(d)
```

Identified,

-   categorical variables: Order, Family, Genus, Species

-   Numerical variables: mass(g), gestation(mo), newborn(g), "weaning(mo), wean mass(g), AFR(mo), max. life(mo), litter size, litters/year, refs

Also, to do a quick exploratory data analysis,

```{r}
skim(d)
```

### Step 01

First, i have to replace all -999 values with NA.

I need to check what this is first, so I could use the skim(d) function above to see what this was actually,

As suggested I am directly using the {naniar} package.

```{r}
d <- d %>%
  replace_with_na_all(condition = ~ .x == -999)

# Also i want to check missingness

miss_var_summary(d)
```

### Step 02

Now I have to ,

-   Drop the variables **litter size** and **refs**.

-   Rename the variable **max. life(mo)** to **maxlife(mo)**.

-   Rename the variable **wean mass(g)** to **weanmass(g)**.

First I need to drop variables "litter size" and refs

```{r}
d <- d %>%
  select(-`litter size`, -refs)

d
```

Now that it is done, I want to do the renamings,

```{r}
d <- d %>%
  rename(
    `maxlife(mo)` = `max. life(mo)`,
    `weanmass(g)` = `wean mass(g)`
  )
d
```

### Step 03
