---
title: "Exercise 11"
format: html
editor: visual
---

## **Practice Model Selection**

First, installing and loading the packages I need.

```{r}
install.packages("naniar")
install.packages("MuMIn")
install.packages("AICcmodavg")
```

```{r}
library(tidyverse)
library(skimr)     
library(naniar)    
library(MuMIn)       
library(AICcmodavg)
```

Loading data,

Since the data file this time is a .txt file, I will use read_tsv() function in R.

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f)

```

For looking at the structure of the data,

```{r}
glimpse(d)
names(d)
```

Identified,

-   categorical variables: Order, Family, Genus, Species

-   Numerical variables: mass(g), gestation(mo), newborn(g), "weaning(mo), wean mass(g), AFR(mo), max. life(mo), litter size, litters/year, refs

Also, to do a quick exploratory data analysis,

```{r}
skim(d)
```

### Step 01

First, i have to replace all -999 values with NA.

I need to check what this is first, so I could use the skim(d) function above to see what this was actually,

As suggested I am directly using the {naniar} package.

```{r}
d <- d %>%
  replace_with_na_all(condition = ~ .x == -999)

# Also i want to check missingness

miss_var_summary(d)
```

### Step 02

Now I have to ,

-   Drop the variables **litter size** and **refs**.

-   Rename the variable **max. life(mo)** to **maxlife(mo)**.

-   Rename the variable **wean mass(g)** to **weanmass(g)**.

First I need to drop variables "litter size" and refs

```{r}
d <- d %>%
  select(-`litter size`, -refs)

d
```

Now that it is done, I want to do the renamings,

```{r}
d <- d %>%
  rename(
    `maxlife(mo)` = `max. life(mo)`,
    `weanmass(g)` = `wean mass(g)`
  )
d
```

### Step 03

Now I have to Log transform all of the other numeric variables,

```{r}
d <- d %>%
  mutate(
    across(
      where(is.numeric),
      ~ log(.x)
    )
  )
d
```

```{r}
summary(d)
```

### Step 04

Here, regressing the guven variables on body mass, we have to perform na.action = na.exclude so that, if some rows have missing values, lm() usually drops them.

1\) Regressing gestation on body mass

```{r}
m_gestation <- lm(`gestation(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

2\) Regressing weaning age on body mass

```{r}
m_weaning <- lm(`weaning(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

3\) Regressing AFR on body mass

```{r}
m_afr <- lm(`AFR(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

4\) Regressing maximum lifespan on body mass

```{r}
m_lifespan <- lm(`maxlife(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

5\) Regressig newborn mass on body mass

```{r}
m_newborn_mass <- lm(`newborn(g)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

6\) Regressing weaning mass on body mass

```{r}
m_weanmass <- lm(`weanmass(g)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

Next, adding the residuals to the dataframe as new variables, I am using resid() function to do this,

```{r}
d <- d %>%
  mutate(
    relGest = resid(m_gestation),
    relWean = resid(m_weaning),
    relAFR = resid(m_afr),
    relLife = resid(m_lifespan),
    relNewbornMass = resid(m_newborn_mass),
    relWeaningMass = resid(m_weanmass)
  )

d
```

### Step 05

01\) Plotting the residuals of max lifespan (relLife) in relation to Order

First, I am going to make the medians of residual values in the descending order

```{r}
life_order_summary <- d %>%
  group_by(order) %>%
  summarise(
    median_relLife = median(relLife, na.rm = TRUE),
    mean_relLife = mean(relLife, na.rm = TRUE),
    n = sum(!is.na(relLife))
  ) %>%
  arrange(desc(median_relLife))

life_order_summary
```

Now I am going to plot the residuals of max lifespan (relLife) in relation to Order, but as the hint says I am going to remove any problematic rows, missing values.

```{r}

d_plot <- d %>%
  filter(!is.na(order), !is.na(relLife)) %>%   
  mutate(
    order = as.factor(order),                
    order = fct_reorder(order, relLife, .fun = median)
  )

```

I have to plot against the log values of max lifespan, because in step 03 we converted all numerical values to their log values.

```{r}
ggplot(d_plot, aes(x = order, y = relLife)) +
  geom_boxplot() +
  
  labs(
    title = "Relative Maximum Lifespan by Mammalian Order",
    x = "Order",
    y = "Residual log(max lifespan)"
  ) +
  theme_bw()
```

This is harder for me to read because the x- axis pints with long names are interfering with each others. I will just flip this in order to read easily.

```{r}
ggplot(d_plot, aes(x = order, y = relLife)) +
  geom_boxplot() +
  coord_flip()
  labs(
    title = "Relative Maximum Lifespan by Mammalian Order",
    x = "Order",
    y = "Residual log(max lifespan)"
  ) +
  theme_bw()
```

Now I can read it easily,

-   From what I can see in my plot, primates have the highest residual lifespan, indicating that species within this order tend to live substantially longer than expected for their body size

-   Also, scandentia, xenartha, carnivora and sirenia also has high residual lifespans.

-   Insectivora, Artiodactyla, and Macroscelidea tend to have lower residual lifespan values, suggesting shorter lifespans relative to body size.

02\) Now plotting residuals of newborn mass (relNewbornMass) in relation to Order.

```{r}
newborn_order_summary <- d %>%
  group_by(order) %>%
  summarise(
    median_relNewbornMass = median(relNewbornMass, na.rm = TRUE),
    mean_relNewbornMass = mean(relNewbornMass, na.rm = TRUE),
    n = sum(!is.na(relNewbornMass))
  ) %>%
  arrange(desc(median_relNewbornMass))
```

```{r}
newborn_order_summary
```

Treating for missing values,

```{r}
d_newborn_plot <- d %>%
  filter(!is.na(order), !is.na(relNewbornMass)) %>%   # remove missing values
  mutate(
    order = as.factor(order),
    order = fct_reorder(order, relNewbornMass, .fun = median)
  )
```

Then plotting

```{r}
ggplot(d_newborn_plot, aes(x = order, y = relNewbornMass)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Relative Newborn Mass by Mammalian Order",
    x = "Order",
    y = "Residual log(newborn mass)"
  ) +
  theme_bw()
```

From what I can see, the highest residual newborn mass belongs to Mascroscelidea. Also Cetacea, Artiodactyla, Hyracoidea, Sirenia have relatively high residual newborn mass. These orders have larger newborns than expected for their body size.

03\) Plotting residuals of weaning mass (relWeaningMass) in relation to Order

```{r}
weaningmass_order_summary <- d %>%
  group_by(order) %>%
  summarise(
    median_relWeaningMass = median(relWeaningMass, na.rm = TRUE),
    mean_relWeaningMass = mean(relWeaningMass, na.rm = TRUE),
    n = sum(!is.na(relWeaningMass))
  ) %>%
  arrange(desc(median_relWeaningMass))

weaningmass_order_summary

```

```{r}
d_weanmass_plot <- d %>%
  filter(!is.na(order), !is.na(relWeaningMass)) %>%   # remove missing values
  mutate(
    order = as.factor(order),
    order = fct_reorder(order, relWeaningMass, .fun = median)
  )

```

```{r}
ggplot(d_weanmass_plot, aes(x = order, y = relWeaningMass)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Relative Weaning Mass by Mammalian Order",
    x = "Order",
    y = "Residual log(weaning mass)"
  ) +
  theme_bw()
```

From what I can see, highest residual weaning mass belongs to Perissodactyla.It is clearly high than other orders.

Catcea, Artidactyla, Primates and scandentia also have high resiidual weaning masses.These orders have larger weaning mass than expected for their body size
