---
title: "Exercise 11"
format: html
editor: visual
---

## **Practice Model Selection**

First, installing and loading the packages I need.

```{r}
# install.packages("naniar")
# install.packages("MuMIn")
# install.packages("AICcmodavg")
```

```{r}
library(tidyverse)
library(skimr)     
library(naniar)    
library(MuMIn)       
library(AICcmodavg)
```

Loading data,

Since the data file this time is a .txt file, I will use read_tsv() function in R.

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f)

```

For looking at the structure of the data,

```{r}
glimpse(d)
names(d)
```

Identified,

-   categorical variables: Order, Family, Genus, Species

-   Numerical variables: mass(g), gestation(mo), newborn(g), "weaning(mo), wean mass(g), AFR(mo), max. life(mo), litter size, litters/year, refs

Also, to do a quick exploratory data analysis,

```{r}
skim(d)
```

### Step 01

First, i have to replace all -999 values with NA.

I need to check what this is first, so I could use the skim(d) function above to see what this was actually,

As suggested I am directly using the {naniar} package.

```{r}
d <- d %>%
  replace_with_na_all(condition = ~ .x == -999)

# Also i want to check missingness

miss_var_summary(d)
```

### Step 02

Now I have to ,

-   Drop the variables **litter size** and **refs**.

-   Rename the variable **max. life(mo)** to **maxlife(mo)**.

-   Rename the variable **wean mass(g)** to **weanmass(g)**.

First I need to drop variables "litter size" and refs

```{r}
d <- d %>%
  select(-`litter size`, -refs)

d
```

Now that it is done, I want to do the renamings,

```{r}
d <- d %>%
  rename(
    `maxlife(mo)` = `max. life(mo)`,
    `weanmass(g)` = `wean mass(g)`
  )
d
```

### Step 03

Now I have to Log transform all of the other numeric variables,

```{r}
d <- d %>%
  mutate(
    across(
      where(is.numeric),
      ~ log(.x)
    )
  )
d
```

```{r}
summary(d)
```

### Step 04

Here, regressing the guven variables on body mass, we have to perform na.action = na.exclude so that, if some rows have missing values, lm() usually drops them.

1\) Regressing gestation on body mass

```{r}
m_gestation <- lm(`gestation(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

2\) Regressing weaning age on body mass

```{r}
m_weaning <- lm(`weaning(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

3\) Regressing AFR on body mass

```{r}
m_afr <- lm(`AFR(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

4\) Regressing maximum lifespan on body mass

```{r}
m_lifespan <- lm(`maxlife(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

5\) Regressig newborn mass on body mass

```{r}
m_newborn_mass <- lm(`newborn(g)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

6\) Regressing weaning mass on body mass

```{r}
m_weanmass <- lm(`weanmass(g)` ~ `mass(g)`, data = d, na.action = na.exclude)
```

Next, adding the residuals to the dataframe as new variables, I am using resid() function to do this,

```{r}
d <- d %>%
  mutate(
    relGest = resid(m_gestation),
    relWean = resid(m_weaning),
    relAFR = resid(m_afr),
    relLife = resid(m_lifespan),
    relNewbornMass = resid(m_newborn_mass),
    relWeaningMass = resid(m_weanmass)
  )

d
```

### Step 05

01\) Plotting the residuals of max lifespan (relLife) in relation to Order

First, I am going to make the medians of residual values in the descending order

```{r}
life_order_summary <- d %>%
  group_by(order) %>%
  summarise(
    median_relLife = median(relLife, na.rm = TRUE),
    mean_relLife = mean(relLife, na.rm = TRUE),
    n = sum(!is.na(relLife))
  ) %>%
  arrange(desc(median_relLife))

life_order_summary
```

Now I am going to plot the residuals of max lifespan (relLife) in relation to Order, but as the hint says I am going to remove any problematic rows, missing values.

```{r}

d_plot <- d %>%
  filter(!is.na(order), !is.na(relLife)) %>%   
  mutate(
    order = as.factor(order),                
    order = fct_reorder(order, relLife, .fun = median)
  )

```

I have to plot against the log values of max lifespan, because in step 03 we converted all numerical values to their log values.

```{r}
ggplot(d_plot, aes(x = order, y = relLife)) +
  geom_boxplot() +
  
  labs(
    title = "Relative Maximum Lifespan by Mammalian Order",
    x = "Order",
    y = "Residual log(max lifespan)"
  ) +
  theme_bw()
```

This is harder for me to read because the x- axis pints with long names are interfering with each others. I will just flip this in order to read easily.

```{r}
ggplot(d_plot, aes(x = order, y = relLife)) +
  geom_boxplot() +
  coord_flip()
  labs(
    title = "Relative Maximum Lifespan by Mammalian Order",
    x = "Order",
    y = "Residual log(max lifespan)"
  ) +
  theme_bw()
```

Now I can read it easily,

-   From what I can see in my plot, primates have the highest residual lifespan, indicating that species within this order tend to live substantially longer than expected for their body size

-   Also, scandentia, xenartha, carnivora and sirenia also has high residual lifespans.

-   Insectivora, Artiodactyla, and Macroscelidea tend to have lower residual lifespan values, suggesting shorter lifespans relative to body size.

02\) Now plotting residuals of newborn mass (relNewbornMass) in relation to Order.

```{r}
newborn_order_summary <- d %>%
  group_by(order) %>%
  summarise(
    median_relNewbornMass = median(relNewbornMass, na.rm = TRUE),
    mean_relNewbornMass = mean(relNewbornMass, na.rm = TRUE),
    n = sum(!is.na(relNewbornMass))
  ) %>%
  arrange(desc(median_relNewbornMass))
```

```{r}
newborn_order_summary
```

Treating for missing values,

```{r}
d_newborn_plot <- d %>%
  filter(!is.na(order), !is.na(relNewbornMass)) %>%   # remove missing values
  mutate(
    order = as.factor(order),
    order = fct_reorder(order, relNewbornMass, .fun = median)
  )
```

Then plotting

```{r}
ggplot(d_newborn_plot, aes(x = order, y = relNewbornMass)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Relative Newborn Mass by Mammalian Order",
    x = "Order",
    y = "Residual log(newborn mass)"
  ) +
  theme_bw()
```

From what I can see, the highest residual newborn mass belongs to Mascroscelidea. Also Cetacea, Artiodactyla, Hyracoidea, Sirenia have relatively high residual newborn mass. These orders have larger newborns than expected for their body size.

03\) Plotting residuals of weaning mass (relWeaningMass) in relation to Order

```{r}
weaningmass_order_summary <- d %>%
  group_by(order) %>%
  summarise(
    median_relWeaningMass = median(relWeaningMass, na.rm = TRUE),
    mean_relWeaningMass = mean(relWeaningMass, na.rm = TRUE),
    n = sum(!is.na(relWeaningMass))
  ) %>%
  arrange(desc(median_relWeaningMass))

weaningmass_order_summary

```

```{r}
d_weanmass_plot <- d %>%
  filter(!is.na(order), !is.na(relWeaningMass)) %>%   # remove missing values
  mutate(
    order = as.factor(order),
    order = fct_reorder(order, relWeaningMass, .fun = median)
  )

```

```{r}
ggplot(d_weanmass_plot, aes(x = order, y = relWeaningMass)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Relative Weaning Mass by Mammalian Order",
    x = "Order",
    y = "Residual log(weaning mass)"
  ) +
  theme_bw()
```

From what I can see, highest residual weaning mass belongs to Perissodactyla.It is clearly high than other orders.

Catcea, Artidactyla, Primates and scandentia also have high resiidual weaning masses.These orders have larger weaning mass than expected for their body size.

### Step 06

Now I have to run models and a model selection process to evaluate what (now log transformed) variables best predict,

01\) maxlife(mo) from the predictors of : gestation(mo), newborn(g), weaning(mo), weanmass(g), litters/year, and overall body mass(g).

02\) AFR(mo) from the predictors of: gestation(mo), newborn(g), weaning(mo), weanmass(g), litters/year, and overall body mass(g).

But, before begining, I should first use drop_na() , because I will need complete cases for response plus all candidate predictors for fair AICc comparison. Every candidate model should use same rows.

01\) I am first going to do this for, maximum lifespan model

```{r}
d_maxlife <- d %>%
  drop_na(
    `maxlife(mo)`,
    `gestation(mo)`,
    `newborn(g)`,
    `weaning(mo)`,
    `weanmass(g)`,
    `litters/year`,
    `mass(g)`
  )
d_maxlife
```

From what we learned at the class last time, now fitting for the global model, and then I can use dredge() in the {MuMIn} package, which explores subsets of agiven global model in an automated way.

To do that, with dredge(), I first have to run aglobal model that includes all of the terms I want to consider and then run the function to fit subsets of that model without having to specify explicitly which submodels I want to run.

So, first fitting for the global model,

```{r}
m_maxlife_global <- lm(
  `maxlife(mo)` ~ `gestation(mo)` + `newborn(g)` + `weaning(mo)` +
    `weanmass(g)` + `litters/year` + `mass(g)`,
  data = d_maxlife,
  na.action = na.fail
)
```

Then runnig all subsets model selection based on AICc.

```{r}
maxlife_dredge <- dredge(m_maxlife_global)
maxlife_dredge
```

Okay, this is hard for me to read, I have to convert this to a readable format, so converting to a data frame

```{r}
maxlife_table <- as.data.frame(maxlife_dredge)
maxlife_table

```

I\) So, finding the best model now, it hsould be at the top of the dredge tabke with delta equal to zero.

```{r}
best_maxlife_model <- get.models(maxlife_dredge, 1)[[1]]
summary(best_maxlife_model)
formula(best_maxlife_model)
```

So, the best model overall based on AICc is

**\`maxlife(mo)\` \~ \`gestation(mo)\` + \`litters/year\` + \`mass(g)\` + \`weaning(mo)\` + 1**

ii\) To find how many models have a delta AICc of 4 or less, I can do the following,

```{r}
top_maxlife_table <- maxlife_table %>%
  filter(delta <= 4)

nrow(top_maxlife_table)

top_maxlife_table
```

So, there are 5 models have a delta AICc of 4 or less.

iii\) To find what variables, if any, appear in all of this set of “top” models, first I will get all the top models,

```{r}
top_maxlife_models <- get.models(maxlife_dredge, subset = delta <= 4)
top_maxlife_models

```

iv\) Now I need to find what predictors appear commonly at every top model above,

```{r}
common_maxlife <- Reduce(
  intersect,
  lapply(top_maxlife_models, function(m) attr(terms(m), "term.labels"))
)

common_maxlife
```

So, \`gestation(mo)\` ,\`litters/year\`and \`weaning(mo)\` are the predictors commonly appearing at every top model.

v\) Next calculating the model average for max life.

```{r}
maxlife_avg <- model.avg(maxlife_dredge, subset = delta <= 4, fit = TRUE)

summary(maxlife_avg)

```

vi\) plotting it,

```{r}
confint(maxlife_avg)

plot(maxlife_avg, full = TRUE)
```

02\) Then doing the pretty much same for the model of AFR(mo) from the predictors of: gestation(mo), newborn(g), weaning(mo), weanmass(g), litters/year, and overall body mass(g).

Here also, before begining, I should first use drop_na() , because I will need complete cases for response plus all candidate predictors for fair AICc comparison. Every candidate model should use same rows.

```{r}
d_afr <- d %>%
  drop_na(
    `AFR(mo)`,
    `gestation(mo)`,
    `newborn(g)`,
    `weaning(mo)`,
    `weanmass(g)`,
    `litters/year`,
    `mass(g)`
  )
d_afr
```

Doing similarly to what I did in the first part, from what we learned at the class last time, now fitting for the global model,

```{r}
m_afr_global <- lm(
  `AFR(mo)` ~ `gestation(mo)` + `newborn(g)` + `weaning(mo)` +
    `weanmass(g)` + `litters/year` + `mass(g)`,
  data = d_afr,
  na.action = na.fail
)
```

Now with dredge(), I will run the function to fit subsets of that model without having to specify explicitly which submodels I want to run.

Then runnig all subsets model selection based on AICc.

```{r}
afr_dredge <- dredge(m_afr_global)
afr_dredge
```

Okay, now too this is hard for me to read, I have to convert this to a readable format, so converting to a data frame

```{r}
afr_table <- as.data.frame(afr_dredge)
afr_table
```

I\) So, finding the best model now, it should be at the top of the dredge tabke with delta equal to zero.

```{r}
best_afr_model <- get.models(afr_dredge, 1)[[1]]
summary(best_afr_model)
formula(best_afr_model)

```

So, the best model overall based on AICc is

**AFR(mo)\` \~ \`gestation(mo)\` + \`litters/year\` + \`mass(g)\` + \`weaning(mo)\` + 1**

ii\) To find how many models have a delta AICc of 4 or less, I can do the following,

```{r}
top_afr_table <- afr_table %>%
  filter(delta <= 4)

nrow(top_afr_table)

top_afr_table
```

So, there are 7 models have a delta AICc of 4 or less.

iii\) To find what variables, if any, appear in all of this set of “top” models, first I will get all the top models,

```{r}
top_afr_models <- get.models(afr_dredge, subset = delta <= 4)
top_afr_models

```

iv\) Now I need to find what predictors appear commonly at every top model above,

```{r}
common_terms_afr <- Reduce(
  intersect,
  lapply(top_afr_models, function(m) attr(terms(m), "term.labels"))
)

common_terms_afr
```

So, \`gestation(mo)\` ,\`litters/year\`and \`weaning(mo)\` are the predictors commonly appearing at every top model.

v\) Next calculating the model average for AFR.

```{r}
afr_avg <- model.avg(afr_dredge, subset = delta <= 4, fit = TRUE)

summary(afr_avg)

```

vi\) plotting it,

```{r}
confint(afr_avg)

plot(afr_avg, full = TRUE)
```
